---
title: "diagnostics"
author: "Nicholas Tierney"
date: "`r format(Sys.time(), '%x at %H:%M:%S %Z')`"
output:
  html_document:
    keep_md: yes
    theme: flatly
    toc: yes
    toc_float: yes
params:
    n_iterations_adapt: 1000
    n_iterations_burnin: 500
    n_iterations_model: 500
    n_thin: 1
    outcome: "fev1_perc"
---

Bleep bloop, I'm running a script to generate these diagnostic plots.

Here is some important information:

- The model predicted `r params$outcome` 
- # adaptive samples was `r params$n_iterations_adapt`
- # burnin samples was `r params$n_iterations_burnin`
- # burnin model iterations was `r params$n_iterations_model`
- # thinning of samples was `r params$n_thin`


```{r setup, include = FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      cache = TRUE,
                      fig.align = "center",
                      fig.show = "hold",
                      autodep = TRUE,
                      dpi = 100)

analysis_path <- "/Users/tierneyn/PhD/code/R/njtbatohs"
knitr::opts_knit$set(analysis_path)

```


```{r load-libraries, include = FALSE}

library(plyr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(purrr)
library(rjags)
library(ggmcmc)
library(ggthemes)
library(ggExtra)
library(data.table)

```


```{r load-output, echo = TRUE, cache.lazy=FALSE}

system.time(
    jags_model_dt <- fread(sprintf("%s/analysis/outputs/03_post_process_output/output_%s/jags_model_dt.csv",
              analysis_path,
              params$outcome))
)

system.time(
jags_model_fit <-
    readRDS(sprintf("%s/analysis/outputs/02_fit_output/output_%s/jags_model_fit.rds",
                    analysis_path,
                    params$outcome))
)


```

# Diagnostics


```{r tidy-model-diagnosis, include = TRUE, cache.lazy = FALSE, echo = TRUE}

library(ggmcmc)

system.time(
jags_diagnosis_plot <- ggs(jags_model_fit)
)

```

```{r list-parameters}
model_params <- c("y_pred", # predictions from Y
                  "mu", # mu results
                  "beta_all", #
                  "beta_d", ## instead of beta[ID[i]], is is beta_d[ID[i]]
                  "beta_0", ## instead of alpha[ID[i]], it is beta_0[ID[i]]
                  "beta_dc", # the overall intercept for the number of days
                  "beta_0c", # the overall intercept for the... intercept.
                  "tau_0", # tau_0 instead of tau_alpha
                  "tau_d", # tau_d instead of tau_beta
                  "tau_y",
                  "sigma_0", # sigma_0 instead of sigma_alpha
                  "sigma_d", # sigma_d instead of sigma_beta
                  "sigma_y")
```


```{r filter-y-pred}

# system.time({
### Y.pred -------------------------------------------------------------------
jags_model_tidy_y_pred_1_5 <- jags_diagnosis_plot %>%
    filter(
        Parameter == "y_pred[1]" |
           Parameter == "y_pred[2]" 
        )

```

```{r filter-mu}

### mu -----------------------------------------------------------------------

jags_model_tidy_mu_1_5 <- jags_diagnosis_plot %>%
    filter(
        Parameter == "mu[1]" |
        Parameter == "mu[2]" 
        )

```

```{r filter-beta-all-1-6}

### beta.all 1 - 6--------------------------------------------------------------
jags_model_tidy_beta_all_1_6 <- jags_diagnosis_plot %>%
    filter(Parameter == "beta_all[1]" |
           Parameter == "beta_all[2]" |
           Parameter == "beta_all[3]" |
           Parameter == "beta_all[4]" |
           Parameter == "beta_all[5]" |
           Parameter == "beta_all[6]" )

```

```{r filter-all-beta-all}

system.time(
jags_model_tidy_all_beta_all <- jags_diagnosis_plot %>%
    filter(grepl("beta_all", Parameter))
)
```


```{r filter-beta-all-7-12}

### beta.all 7 - 12-------------------------------------------------------------
jags_model_tidy_beta_all_7_12 <- jags_diagnosis_plot %>%
    filter(Parameter == "beta_all[7]" |
           Parameter == "beta_all[8]" |
           Parameter == "beta_all[9]" |
           Parameter == "beta_all[10]" |
           Parameter == "beta_all[11]" |
           Parameter == "beta_all[12]" )

```

```{r filter-beta-all-13-17}

### beta.all 13 - 17------------------------------------------------------------
jags_model_tidy_beta_all_13_17 <- jags_diagnosis_plot %>%
    filter(Parameter == "beta_all[13]" |
           Parameter == "beta_all[14]" |
           Parameter == "beta_all[15]" |
           Parameter == "beta_all[16]" |
           Parameter == "beta_all[17]" )

```

```{r filter-beta-d-1-5}

### beta_d ---------------------------------------------------------------------
jags_model_tidy_beta_d_1_5 <- jags_diagnosis_plot %>%
    filter(Parameter == "beta_d[1]" |
           Parameter == "beta_d[2]" 
           )

```

```{r filter-beta-0-1-5}

### beta_0 ---------------------------------------------------------------------
jags_model_tidy_beta_0_1_5 <- jags_diagnosis_plot %>%
    filter(Parameter == "beta_0[1]" |
           Parameter == "beta_0[2]" 
           )

```

```{r filter-beta-dc}

### beta_dc ---------------------------------------------------------------------
# the overall intercept for the number of days
jags_model_tidy_beta_dc <- jags_diagnosis_plot %>%
    filter(Parameter == "beta_dc")

```

```{r filter-beta-0c}

### beta_dc ---------------------------------------------------------------------
# the overall intercept for the... intercept.
jags_model_tidy_beta_0c <- jags_diagnosis_plot %>%
    filter(Parameter == "beta_0c")

```

```{r filter-tau-sigma}

### tau ----------------------------------------------------------------------
jags_model_tidy_taus <- jags_diagnosis_plot %>%
    filter(Parameter == "tau_0" | # tau_0 instead of tau_alpha
           Parameter == "tau_d" | # tau_d instead of tau_beta
           Parameter == "tau_y")


### sigma --------------------------------------------------------------------
jags_model_tidy_sigmas <- jags_diagnosis_plot %>%
    filter(Parameter == "sigma_0" | # sigma_0 instead of tau_alpha
           Parameter == "sigma_d" | # sigma_d instead of tau_beta
           Parameter == "sigma_y")
# })
```


```{r fun-trace-dens}
# this function places the traceplots and densities side by side
ggs_trace_dens <- function(x){
    
p1 <- ggs_traceplot(x) + theme(legend.position = "none")
p2 <- ggs_density(x) + theme(legend.position = "none")

gridExtra::grid.arrange(p1,
                        p2,
                        ncol = 2, 
                        nrow = 1)

}

```

## Outcome


```{r diagnostics-outcome, fig.height = 10, fig.width=8, fig.asp = NULL, out.width = "100%", fig.align = "default"}

# density and traceplot
ggs_trace_dens(jags_model_tidy_y_pred_1_5)

# })
```

# ### mu 

```{r diagnostics-mu, fig.height = 10, fig.width=8, fig.asp = NULL, out.width = "100%", fig.align = "default"}

# system.time({
# density and traceplot
ggs_trace_dens(jags_model_tidy_mu_1_5)
# })
```

## All beta terms

```{r diagnostics-all-beta-all, fig.height = 30, fig.width=8, fig.asp = NULL, out.width = "100%", fig.align = "default"}
# 
# system.time({
    
# density and traceplot
ggs_trace_dens(jags_model_tidy_all_beta_all)
# })


## Alpha (individual Intercept)-----------------------------------------------

```{r diagnostics-alpha-i, fig.height = 10, fig.width=8, fig.asp = NULL, out.width = "100%", fig.align = "default"}

# 
# system.time({
    
# density and traceplot
ggs_trace_dens(jags_model_tidy_beta_0_1_5)

# })
```


## Beta (Individual change from population)

```{r diagnostics-beta-i, fig.height = 10, fig.width=8, fig.asp = NULL, out.width = "100%", fig.align = "default"}
# 
# system.time({
    
# density and traceplot
ggs_trace_dens(jags_model_tidy_beta_d_1_5)


# })
```

## taus 

```{r diagnostics-taus, fig.height = 10, fig.width=8, fig.asp = NULL, out.width = "100%", fig.align = "default"}

# system.time({
    
# density and traceplot
ggs_trace_dens(jags_model_tidy_taus)



# })
```

## Sigmas 


```{r diagnostics-sigmas, fig.height = 10, fig.width=8, fig.asp = NULL, out.width = "100%", fig.align = "default"}

# system.time({
    
# density and traceplot
ggs_trace_dens(jags_model_tidy_sigmas)



# })
```

# Autocorrelation plots

```{r autocorrelation-plots, fig.height = 10, fig.width=8, fig.asp = NULL, out.width = "100%", fig.align = "default"}
## Autocorrelation
## Y.pred
ggs_autocorrelation(jags_model_tidy_y_pred_1_5)

ggs_autocorrelation(jags_model_tidy_mu_1_5)

ggs_autocorrelation(jags_model_tidy_all_beta_all)

ggs_autocorrelation(jags_model_tidy_beta_all_1_6)

ggs_autocorrelation(jags_model_tidy_beta_all_7_12)

ggs_autocorrelation(jags_model_tidy_beta_all_13_17)

ggs_autocorrelation(jags_model_tidy_beta_0_1_5)

ggs_autocorrelation(jags_model_tidy_beta_d_1_5)

ggs_autocorrelation(jags_model_tidy_taus)

ggs_autocorrelation(jags_model_tidy_sigmas)
```

# Further diagnostics


```{r}

jags_model_fit_bmi <- jags_model_fit_all %>% pluck(1) %>% pluck(1)
jags_model_fit_dias <- jags_model_fit_all %>% pluck(2) %>% pluck(1)
jags_model_fit_sys <- jags_model_fit_all %>% pluck(3) %>% pluck(1)
jags_model_fit_fev1 <- jags_model_fit_all %>% pluck(4) %>% pluck(1)

all_jags_fits <- list(
  bmi = jags_model_fit_bmi,
  dias = jags_model_fit_dias,
  sys = jags_model_fit_sys,
  fev1 = jags_model_fit_fev1
)

system.time(
all_ess <- map(all_jags_fits, coda::effectiveSize)
)

system.time(
ess <- coda::effectiveSize(jags_model_fit[[1]])
)
system.time(
geweke_diag <- coda::geweke.diag(jags_model_fit[[1]])
)

geweke_df <- tibble(parameter = names(geweke_diag$z),
                    geweke = geweke_diag$z)

ess_df <- tibble(parameter = names(ess),
                 ess = ess)


geweke_df %>%
  ggplot(aes(x = geweke)) + 
  geom_density()

ess_df %>%
  ggplot(aes(x = ess)) + 
  geom_density()

```
