---
title: "Bayesian modelling to assist inference on health outcomes in Occupational Health Surveillance"
author: Nicholas J Tierney(1,2,3), Samuel Clifford(1,2), Christopher C Drovandi(1,2),
  Kerrie L Mengersen(1,2)
output:
  pdf_document:
    includes:
      in_header: premble.tex
    keep_tex: yes
  html_document:
    keep_md: yes
  word_document:
    reference_docx: paper-template.docx
params:
  data: 2015
  n_iterations_adapt: 1000
  n_iterations_burnin: 10000
  n_iterations_model: 20000
  n_thin: 1
bibliography: paper.bib
csl: vancouver.csl
---

Affiliations:

1. ARC Centre of Excellence for Mathematical and Statistical Frontiers 

2. School of Mathematical Sciences, Queensland University of Technology, 
Brisbane, Queensland, Australia 

3. Department of Econometrics and Business Statistics, Monash University, 
Melbourne, Victoria, Australia


**Corresponding Author: Nicholas J Tierney**

**Corresponding Author Address: Department of Econometrics and Business Statistics E774, Menzies Building Monash University, Clayton, Victoria, Australia, 3800**

**Corresponding Author email: nicholas.tierney@monash.edu**

**Corresponding Author phone: 03 9905 9352**


```{r setup, include = FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      cache = TRUE,
                      dev = c("png"),
                      autodep = TRUE,
                      dpi = 600)

# use rproj root to find the rproject root file
knitr::opts_knit$set(rprojroot::find_rstudio_root_file())

analysis_path <- rprojroot::find_rstudio_root_file()
```

```{r load-libraries, include = FALSE}
library(dplyr)
library(tidyr)
library(forcats)
library(hms)
library(modelr)
library(stringr)
library(lubridate)
library(purrr)
library(rjags)
library(ggmcmc)
library(ggthemes)
library(readr)
library(readxl)
library(lubridate)
library(tibble)
library(data.table)
library(broom)
```


```{r source-plotting-functions}
source(sprintf("%s/R/create_obs_pred_summary.R", analysis_path))
source(sprintf("%s/R/plot_fig_four.R", analysis_path))
```

```{r read-03-post-process-output-fev1, include = FALSE}

system.time(
jags_model_dt_summary_fev1 <- fread(
sprintf("%s/analysis/outputs/03_post_process_output/output_fev1_perc/jags_model_dt_summary.csv",
         analysis_path)) %>%
    mutate(model_outcome = "fev1") %>%
    as.data.table()
)

system.time(
dat_isba_fev1 <- feather::read_feather(
sprintf("%s/analysis/outputs/01_tidy_output/output_fev1_perc/dat_isba.feather",
         analysis_path)
)
)

system.time(
jags_dat_model_fev1 <- readRDS(
sprintf("%s/analysis/outputs/01_tidy_output/output_fev1_perc/jags_dat_model.rds",
         analysis_path)
)
)

system.time(
jags_dat_prep_fev1 <- readRDS(
    sprintf("%s/analysis/outputs/01_tidy_output/output_fev1_perc/jags_dat_prep.rds",
         analysis_path)
    )
)

```

```{r read-03-post-process-output-bmi, include = FALSE}
system.time(
jags_model_dt_summary_bmi <- fread(
sprintf("%s/analysis/outputs/03_post_process_output/output_bmi/jags_model_dt_summary.csv",
         analysis_path)
) %>%
    mutate(model_outcome = "bmi") %>%
    as.data.table()
)

system.time(
dat_isba_bmi <- feather::read_feather(
sprintf("%s/analysis/outputs/01_tidy_output/output_bmi/dat_isba.feather",
         analysis_path)
)
)

system.time(
jags_dat_model_bmi <- readRDS(
sprintf("%s/analysis/outputs/01_tidy_output/output_bmi/jags_dat_model.rds",
         analysis_path)
)
)

system.time(
jags_dat_prep_bmi <- readRDS(
    sprintf("%s/analysis/outputs/01_tidy_output/output_bmi/jags_dat_prep.rds",
         analysis_path)
    )
)


```

```{r read-03-post-process-output-sys, include = FALSE}

system.time(
jags_model_dt_summary_sys <- fread(
sprintf("%s/analysis/outputs/03_post_process_output/output_sys/jags_model_dt_summary.csv",
         analysis_path)
) %>%
    mutate(model_outcome = "sys") %>%
    as.data.table()
)

system.time(
dat_isba_sys <- feather::read_feather(
sprintf("%s/analysis/outputs/01_tidy_output/output_sys/dat_isba.feather",
         analysis_path)
)
)

system.time(
jags_dat_model_sys <- readRDS(
sprintf("%s/analysis/outputs/01_tidy_output/output_sys/jags_dat_model.rds",
         analysis_path)
)
)

system.time(
jags_dat_prep_sys <- readRDS(
    sprintf("%s/analysis/outputs/01_tidy_output/output_sys/jags_dat_prep.rds",
         analysis_path)
    )
)


```

```{r read-03-post-process-output-dias, include = FALSE}
system.time(
jags_model_dt_summary_dias <- fread(
sprintf("%s/analysis/outputs/03_post_process_output/output_dias/jags_model_dt_summary.csv",
         analysis_path)
) %>%
    mutate(model_outcome = "dias") %>%
    as.data.table()
)

system.time(
dat_isba_dias <- feather::read_feather(
sprintf("%s/analysis/outputs/01_tidy_output/output_dias/dat_isba.feather",
         analysis_path)
)
)

system.time(
jags_dat_model_dias <- readRDS(
sprintf("%s/analysis/outputs/01_tidy_output/output_dias/jags_dat_model.rds",
         analysis_path)
)
)

system.time(
jags_dat_prep_dias <- readRDS(
    sprintf("%s/analysis/outputs/01_tidy_output/output_dias/jags_dat_prep.rds",
         analysis_path)
    )
)


```


## Abstract

**Objectives**

Occupational Health Surveillance (OHS) facilitates early detection of disease
and dangerous exposures in the workplace. Current OHS analyses ignore important
workplace structures and repeated measurements. There is
a need to provide systematic analyses of medical data that incorporate the data
structure. Although multilevel statistical models may account for features of
OHS data, current applications in occupational health medicine are often not
appropriate for OHS. Additionally, typical OHS data has not been analysed in a
Bayesian framework, which allows for calculation of probabilities of potential
events and outcomes. This paper's objective is to illustrate the use of
Bayesian modeling of OHS. Three analytic aims are addressed: (1) Identify patterns and changes in health outcomes; (2) Explore the effects of a particular risk factor, smoking and industrial exposures over time for individuals and worker groups; (3) identify risk of chronic conditions in individuals.

**Method**

A Bayesian hierarchical model was developed to provide individual and group
level estimates and inferences for health outcomes, FEV1%, BMI, and
Diastolic and Systolic blood pressure.

**Results**

We identified individuals with the greatest degree of change over time for each
outcome, and demonstrated how to flag individuals with substantive negative
health outcome change. We also assigned probabilities of individuals moving into "at risk" health categories one year from their last visit.

**Conclusion**

Bayesian models can account for features typically encountered in OHS data,
such as individual repeated measurements and group structures. We
describe one way to fit these data and obtain informative estimates and
predictions of employee health.

## Summary

**Strengths and Limitations of the study**

- Strength: This is the first application of Bayesian methods to typical data
found in occupational health surveillance.

- Strength: The methods used account for important features
of data such as multiple measurements for individuals, and the group structure
of exposure groups in the workplace.

- Strength: The model allows for groups and individuals to be flagged as "at risk", enabling proactive action on individual health.

- Strength: The definition of a Bayesian credible interval as a range of
probable values for a parameter makes it easier to communicate model inferences.

- Strength: The model provides probabilities of interest directly, conditional on the data, which is a useful complement to credible intervals that makes effects and uncertainty simpler and easier to communicate to health practitioners.

- Limitation: No account was taken of the healthy worker effect, so whilst the focus of the paper is on employees rather than the
general population, the analysis may be biased if healthier employees remain
longer in the industry. 

- Limitation: The model used vague priors, and so future work could explore
the use of more informative priors, based on, for example, previous data collected in similar fields. 

## Data Sharing

For reasons of employee confidentiality and industry sensitivity, the dataset
used in the case study is not available for sharing. However, the seminal
features of the dataset are well described in the paper.

# Contributors 

NJT conducted literature review, statistical analysis, created visualisations, and wrote the first draft. SC provided critical feedback, assisted in developing the statistical model and in presenting results. CCD provided assistance in developing the statistical model and in critical feedback of the paper. KLM provided initial description of the model, assisted in developing the code, and provided critical feedback of the paper and presentation of results. All authors approved the version of the paper for publishing, and agreed to respond to questions that may arise regarding integrity of the work.

# Funding 

This research was jointly funded by an Australian Postgraduate Award (APA), the Australian Technology Network Industry Doctoral Training Centre (IDTC), the Australian Research Council, and the ARC Centre of Excellence for Mathematical and Statistical Frontiers. CCD was supported by an Australian Research Council’s Discovery Early Career Researcher Award funding scheme (DE160100741).

# Competing interests 

None declared.

# Ethics approval 

The QUT University Human Research Ethics Committee.

# Introduction

Occupational Health Surveillance (OHS) is the systematic collection, analysis,
and dissemination of employee exposure and health data to facilitate early
detection of disease and dangerous exposures in the workplace [@schulte2008].
Australian employers have a responsibility to identify, assess, and control
risks arising from workplace hazards [@sw_oz; @Firth2006]. There is a rigorous
methodology for OHS data collection, but a surprising lack of agreement about
analysis of these data [@lewis2013]. Indeed, industry OHS data collection is
often targeted for managing risk and implementing engineering controls.
Consequently, many of the analyses conducted in industry focus on the likelihood
of exposure rather than the impact of these risk factors on health. Moreover,
current practices may ignore important data structures such as repeated
measurements and workplace structures. This results in inferences not being
applicable for individuals over time, or for groups with similar exposures
within the workplace.

There is a need to provide systematic analyses of medical data that incorporate
workplace structure, relevant to risk factors. An example of such a workplace
structure is segmentation of the workplace into similar exposure groups (e.g.,
as in @Banerjee2014). Moreover, such analyses need to incorporate typical
features of OHS data, in particular where individuals have multiple health
measurements, or single repeated measurements over time, or missing data. These
analyses should provide both individual and group health predictions, and should
improve the understanding of exposure effects on the workplace population as a
whole, as well as similar exposure groups and individuals. Such analyses
could flag individuals and groups for further health monitoring. The absence of
such analyses in industry means that chronic disease and dangerous work
environments may go unidentified and that health funding is not optimally or
effectively targeted.

These features of OHS data described above can be accounted for with multilevel
statistical models. These are in wide use in epidemiology [@lynch2005; @duncan1998], and have been used in occupational health medicine to evaluate
decline in lung function for ceramic fibre workers [@McKay2011], assess impacts
of asbestos [@Algranti2013], measure decline from cystic fibrosis
[@Szczesniak2013] and model leptospirosis in abattoir employees [@Cook2016].

However, applications of multilevel models in occupational health medicine do
not quite mimic the analyses conducted in standard industry environments, as
they might ignore individuals with only one measurement, population minority
groups, or workplace structures [@McKay2011; @Algranti2013]. This is likely due
to the fact that the goal of these papers is often to demonstrate the use of a
new method [@deVocht2009; McKay2011; @Algranti2013; @Szczesniak2013], or
discover new health risk factors or exposures [@Bakke1991; @Gibbs2007; @Hellgren2002; @Radon2008; @Olsson2010].

In contrast, the goal of OHS analysis is to provide individual predictions for
health, understand the effect of exposures on groups, and monitor exposures and
health over time, so that individuals and groups at risk of some disease can be
flagged for further health monitoring. Thus, ignoring cases with only one
medical visit or analysing only subsets of the population for reasons such as
sufficient sample size, can increase bias and/or variance of estimates.

Bayesian models provide a pragmatic framework for this research problem, as they
provide simple and effective ways of analysing small effects, and provide a rich
set of results that can be interpreted with probabilistic statements. Bayesian
methodology also allows for direct comparison between groups and individuals,
and provides probabilities on potential events and outcomes.

Bayesian techniques have been recently applied in occupational health, with
@Banerjee2014 demonstrating the use of hierarchical models to combine
monitoring data and professional judgement from occupational hygienists to
facilitate decision making. Bayesian hierarchical models have also been applied
to quantify chemical exposure variation in human populations [@Shao2016], and
to combine two data sources from animal studies and human
industrial studies to create informative priors to estimate human lung function
changes [@Bartell2017].

Industry data used in the literature typically consist of longitudinal data
collected from employees in a particular industry, or set of industries. These
data are used to evaluate the effect of the working conditions, such as long
hours with no sleep [@Sieber2014; @Fekedulegn2016], metal smelting, and other
exposures [@deVocht2009; @Olsson2010; @McKay2011; @Algranti2013; @Szczesniak2013; @Cook2016]. Other OHS studies focus on small populations, using
experiments to evaluate effects of increasing some exposure on health
[@Hendrick1988; @Stenton1990], or larger cross-sectional studies using registry
data, cohort studies, or surveys to evaluate the effect of an environmental
exposure on diseases such as rhinitis or cystic fibrosis [@Szczesniak2013; @Radon2008].

Notwithstanding these studies, there are no examples of Bayesian hierarchical
modelling and analysis of typical OHS data with
applications in an industry context. This paper analyses OHS data from selected
industrial sites around Australia to identify risk factors for health outcomes.
The multilevel model adopted is a Bayesian hierarchical model providing
individual and group level estimates and inferences.

The aims of the analysis are three-fold, and are focused on the following
health outcomes: lung function (as a percentage of predicted Forced Expiratory
Volume in 1 second (FEV~1~, FEV~1~\%), Body Mass Index (BMI), and systolic
and diastolic blood pressure. These health outcomes were selected as they are
clinically substantive in the case study population and are well known in OHS.
The first analytic aim is to identify patterns and changes in health outcomes. The second
aim is to explore the effects of a particular risk factor: smoking and
industrial exposures over time for individuals and worker groups. The third
aim is to identify risk of chronic conditions (such as obesity, hypertension, and
obstructive / restrictive lung disease) in individuals.

# Method

**Case Study Data**

The case study considered in this paper is typical of many large companies that 
are involved in a range of activities, such as construction, mining, manufacturing and agriculture. For reasons of confidentiality, the particular industry and associated sites are not named here. The data are comprised of over 3000 employee medical records from nearly 2000 individuals located at a number of sites. Each observation is a medical visit, and while most employees have one or two visits, some have over ten visits over a 10 year period. Employees are typically grouped by their workplace exposure; for example,
Administration employees are less likely to be
exposed to environmental factors such as dust or noise, compared to maintenance exposure groups. In this way, Administration provides a useful control group to compare to the other exposure groups. Employees may change positions within the company over their career and thus may also change their exposure group. The pattern of measurements over time for
individuals are illustrated in Figure 1, which displays individual measurements
of lung function (FEV1\%) over visits for selected exposure groups.

```{r fig-one, fig.width = 8, fig.height = 5, fig.cap = "Individual employee lung function (FEV1%) over their medical visit number (1, up to 10), for each exposure group. A sample of 50% of employees is used to reduce overplotting.  Individuals are linked by a line between observations. A point on the 2nd or later visit which is not joined to previous points by a line indicates individuals who have changed exposure group. Broken lines and individual floating points without lines indicate where individuals have changed exposure group. Note that the number of days between visits varies by individual and exposure group."}

# choose a random selection of IDs
set.seed(2017-11-30)
sample_id <- sample(unique(dat_isba_fev1$ID),900)

dat_isba_fev1 %>%
    filter(ID %in% sample_id) %>%
    filter(seg_label != "unknown") %>%
    ggplot(data = .,
           aes(x = lf_visit,
               y = fev1_perc,
               group = uin)) +
    geom_line(alpha = 0.8,
              colour = "#003366") +
    geom_point(alpha = 0.8,
               colour = "orange2") +
    facet_wrap(~ seg_label,
               ncol = 3,
               labeller = seg_labeller) +
    theme(legend.position = "none",
          strip.background = element_rect(fill="#003366"),
          strip.text.x = element_text(colour = "white"),
          axis.text.x = element_text(size = 16,
                                     colour = "#003366"),
          axis.text.y = element_text(size = 16,
                                     colour = "#003366"),
          axis.title.y = element_text(size = 16,
                                      colour = "#003366")) +
    labs(x = "Visit #",
         y = "FEV1 %") + 
    theme_hc()


dat_days_sum <-    
dat_isba_bmi %>%
    select(days_since_arrival) %>%
    filter(days_since_arrival != 0) %>%
    summarise(median = median(days_since_arrival),
              lower_025 = quantile(days_since_arrival, probs = 0.025),
               upper_975 = quantile(days_since_arrival, probs = 0.975)) %>%
    round()

```

The frequency of medical visits changes for each exposure group, as
certain exposure groups require more frequent medical examinations to
ensure that they are fit for work. The times between visits for each
worker were not equally spaced, with median number of days since first visit
being `r dat_days_sum$median[1]` (IQR = `r dat_days_sum$lower_025[1]` -
`r dat_days_sum$upper_975[1]`), or `r round(dat_days_sum$median[1]/365,1)` years (IQR = `r round(dat_days_sum$lower_025[1]/365,1)` -
`r round(dat_days_sum$upper_975[1]/365,1)` years). Gender (male or female) and smoking status (ever
smoker or never smoker) were also recorded.  Dust data were not recorded for some dates and
were interpolated using a loess model [@steinle2015] fitted for each exposure
group, so that the values corresponded to medical examination dates. Interpolated values should be treated with care, and explored with visual and numerical summaries.

**Ethics**

The Queensland University of Technology Human Research Ethics Committee assessed
that this research met the conditions for exemption from HREC review and
approval in accordance with section 5.1.22 of the Australian National Statement
on Ethical Conduct in Human Research. 

**Patient and Public Involvement**

The development of the research questions and outcomes were informed by discussion with health practitioners who helped collect the data. The patients were not involved in the results, design, or recruitment. The paper will be shared with the medical practitioners for their use in future designs. We thank the health practitioners and patients involved in the data collection.

**Modelling**

We construct four multilevel Bayesian hierarchical models. Each model predicts
one of the four outcomes: lung function (FEV1\%), Body Mass Index (BMI), systolic
blood pressure, and diastolic blood pressure. 

Let $Y_{ij}$ be the $i^{th}$ individual's $j^{th}$ health observation, at a time
day$_{ij}$ after their first visit. We assume that $Y_{ij}$ follows a normal distribution with mean $\mu_{ij}$ and variance $\sigma_y^2$. Let $\beta_{0i}$ 
and $\beta_{di}$ be respectively the individual intercept and individual health trend coefficient 
associated with the $j^{th}$ day for the $i^{th}$ person; these individual parameters are centered
around an overall intercept $\beta_{0c}$ and an overall slope $\beta_{dc}$, 
the effect of the number of days since arriving at the workplace. Thus
$\beta_{di}$ is the linear
trend over time for the health characteristics of interest for the $i^{th}$ individual, over and above the
overall population effect. Let $\beta_{g}$ be the effect of being female (compared to being male); let $\beta_{s}$ be the effect of being a smoker (compared to a never smoker), and let $\sum_{k = 1}^{n_{\mathrm{exposure}}-1} \beta_k I(\mathrm{exposure}_{ij} = k)$ be the effect of a workplace exposure, where $I(.)$ indicates whether an individual $i$ at a visit $j$ is in exposure group $k$, with the baseline exposure group set to Administration. Thus the model for a particular health outcome is represented as:

$$
Y_{ij} \sim \mathcal{N}(\mu_{ij},\sigma^2_y)
$$

with 
$$
\mu_{ij}=\beta_{0i}+\beta_{di}\mathrm{day}_{ij}+\beta_g\mathrm{gender}_{ij}+\beta_{s}\mathrm{smoke}_{ij}+\beta_p\mathrm{dust}_{ij}+\sum_{k=1}^{n_{\mathrm{exposure}}-1}{\beta_k }I(\mathrm{exposure_{ij}}=k) 
$$

$$
\beta_{0i} \sim N(\beta_{0c},\sigma_{0}^2)
$$

$$
\beta_{di} \sim N(\beta_{dc},\sigma_{d}^2)
$$


for $i=1 \ldots n_{I}$, $j=1 \ldots n_{0i}$, $k=1 \ldots,n_E$, where $n_I$ is the total number of individuals, $n_{0i}$ is the number of observations for each individual, and $n_E$ is the number of exposure groups.

In the absence of other information, all of the regression coefficients were allocated independent normal priors with a mean of 0 and a variance of $D_1 = 10^3$.

$$
\beta_{0c},\beta_{dc},\beta_{di},\beta_g,\beta_s,\beta_p,\beta_k \sim N(0,D_1)
$$

Priors on $\sigma_{0},\sigma_y,\sigma_{d}$ were set to a uniform distribution
with bounds of zero and $D2$, where $D_2=100$ for BMI and FEV~1~\%, and $D_2=50$ for Systolic and Diastolic blood pressure. $D_2$ is intended to better reflect the variation in BMI and FEV~1~% compared to blood pressure. Note also that we do not recommend automatically choosing set values for the uniform, but to instead choose sensible bounds based on the problem at hand.

$$
\sigma_{y},\sigma_0,\sigma_{d}, \sim \mathrm{Uniform}(0,D_2)
$$

Note also that the priors used for the $\beta$ terms are proper priors, which produce a proper posterior. In some cases improper priors such as an infinite uniform prior might be used, but these are sometimes not valid choices (See @Hobert1996 and @Rubio2018 for more details). It is worthwhile to consider the choice of prior for the variance terms. Although we have used inverse gamma and uniform priors, other weakly informative priors could be considered, such as a half-t-prior (represented as a half-Cauchy) [@Gelman2006]. It is important to not automatically choose uniform or half-t-priors, but to explore options during model building.

Data processing and manipulation were implemented using the R statistical
programming language [@rcore] and various R packages [@dplyr; @tidyr; @readxl;
@purrr; @lubridate; @data.table]. To ensure reproducibility, the paper
was written using rmarkdown and knitr [@rmarkdown; @knitr].  Potential outliers in the data were checked administratively and confirmed for biological plausibility in the context of the workforce under consideration. Given this, we elected to include them in the analyses. Moreover, the modelling goal is to identify those who are risk, so removing outliers seems counter to that goal. The model was run for 20,000 iterations (10,000
burnin) using JAGS [@plummer2003; @rjags]. Thinning was applied to the analysis, removing every 20th value to assist in reducing autocorrelation and for computational storage. We note that thinning is not absolutely necessary in an analysis, and should be assessed case by case [@Link2012]. We note that other software such as STAN, WinBUGS or OpenBUGS, Nimble, and greta could also have been used [@rstan; @Lunn2000; @nimble; @greta]. The diagnostics for MCMC convergence were predominantly graphical and statistical [@Lunn2012; @ggmcmc]. Graphical evaluation included expert examination of posterior density plots, traceplots and autocorrelation plots of parameters. Statistical evaluation included calculation of the Geweke diagnostic and effective sample size. 

Missing values were imputed from their respective posterior conditional
distributions as part of the Bayesian analysis. Posterior estimates of each parameter, including mean, 95\% and 80\% credible intervals, and probability of being negative
were calculated after burnin. An effect was nominated as substantive if
the corresponding credible interval did not contain zero. 
The probability of individual health outcomes reaching the threshold value of being a chronic condition was also calculated. Individuals were identified as being "at risk" if the corresponding estimates of the parameter for change over time, $\beta_{di}$, contained 0 in the 95\% credible intervals, and $\beta_{di}$ was far away from zero. For the purposes of exposition, individuals with 3 or more visits were selected as examples to explore further.

Patterns and trends in health outcomes were examined by exploring individuals'
change over time and identifying substantive effects, addressing analytic aim 1. The
effects of smoking and industrial exposures over time for individuals and
exposure groups were examined by evaluating substantive effects of smoking and
dust for each outcome, and finding those exposure groups substantively different
from the Administration population, addressing aim 2. To identify future risk
of chronic conditions, one year forecasts for each individual and corresponding
95\% credible intervals were calculated from the respective posterior predictive distribution, and the probability
of having a chronic condition in one year was obtained, addressing aim 3. Model
fit was evaluated by examining the proportion of observed values lying within
the 95\% and 80\% posterior predictive intervals [@BDA3 chapter 6; @Besag1995].

# Results

```{r descriptive-stats}

age_info_gender <- 
dat_isba_bmi %>%
    group_by(uin) %>%
    mutate(max_vis = max(lf_visit)) %>%
    ungroup() %>%
    filter(max_vis == lf_visit) %>%
    mutate(age =  as.numeric((date - dob)) / 365) %>%
    group_by(sex) %>%
    summarise(mean = mean(age),
              sd = sd(age, na.rm = TRUE))

age_info <- 
dat_isba_bmi %>%
    group_by(uin) %>%
    mutate(max_vis = max(lf_visit)) %>%
    ungroup() %>%
    filter(max_vis == lf_visit) %>%
    mutate(age =  as.numeric((date - dob)) / 365) %>%
    summarise(mean = mean(age),
              sd = sd(age))

mean_age <- age_info$mean
sd_age <- age_info$sd

mean_age_female <- age_info_gender %>% filter(sex == "female") %>% .$mean
mean_age_male <- age_info_gender %>% filter(sex == "male") %>% .$mean
sd_age_female <- age_info_gender %>% filter(sex == "female") %>% .$sd
sd_age_male <- age_info_gender %>% filter(sex == "male") %>% .$sd

```


**Demographics**

In the case study dataset, the population was predominantly male (86\%), with the mean overall
age being `r round(mean_age,2)` years. Males were older on average, but not significantly so, compared with females. For all exposure groups there were more males than females, except in the Administration exposure group. The proportion of individuals in selected exposure groups is shown below in Table 1.

```{r new-table-of-segs, results = "asis"}

new_exposure_group <- c("Administration",
                        "Emergency",
                        "Exterior Maintenance",
                        "Technicians",
                        "Technology",
                        "Interior Maintenance",
                        "Field Experts")

table_one_percent_pop_in_exposure <- dat_isba_bmi %>%
    count(seg_label) %>%
    mutate(pct = round((n / sum(n))*100,2)) %>%
    mutate(seg_label = my_seg_case_when(seg_label)) %>%
    select(seg_label,
           pct) %>%
    filter(seg_label %in% new_exposure_group) %>%
    arrange(-pct) %>%
    mutate(pct = round(pct,1)) %>%
    rename(`Exposure` = seg_label,
           `% of Population` = pct) %>%
    # pander::pandoc.table(caption = "Percent of the population in selected exposures")
    knitr::kable(caption = "Percent of the population in selected exposures", )

# table_one_percent_pop_in_exposure
```
|Exposure             | % of Population|
|:--------------------|---------------:|
|Technology           |            18 - 20|
|Administration       |             8 - 10|
|Interior Maintenance |             8 - 10|
|Technicians          |             6 - 7|
|Emergency            |             5 - 6|
|Exterior Maintenance |             4 - 5|
|Field Experts        |             2 - 3|


Table: Percent of the population in selected exposures



```{r test-table-fig-three, fig.out = "70%", fig.width = 10, fig.height = 10}
  
obs_pred_summary_fev1 <- create_obs_pred_summary(jags_model_dt_summary_fev1,
                                                 jags_dat_model_fev1,
                                                 jags_dat_prep_fev1) %>%
                         mutate(model_outcome = "fev1")

obs_pred_summary_bmi <- create_obs_pred_summary(jags_model_dt_summary_bmi,
                                                 jags_dat_model_bmi,
                                                 jags_dat_prep_bmi) %>%
                         mutate(model_outcome = "bmi")

obs_pred_summary_sys <- create_obs_pred_summary(jags_model_dt_summary_sys,
                                                 jags_dat_model_sys,
                                                 jags_dat_prep_sys) %>%
                         mutate(model_outcome = "sys")

obs_pred_summary_dias <- create_obs_pred_summary(jags_model_dt_summary_dias,
                                                 jags_dat_model_dias,
                                                 jags_dat_prep_dias) %>%
                         mutate(model_outcome = "dias")

obs_pred_summary_all <- bind_rows(obs_pred_summary_fev1,
                                  obs_pred_summary_bmi,
                                  obs_pred_summary_sys,
                                  obs_pred_summary_dias)

```

```{r make-rss}

model_rss <- obs_pred_summary_all %>%
    filter(forecasted == FALSE) %>%
    # group_by(model_outcome) %>%
    mutate(resid = (outcome - pred_mean)^2) %>% 
    group_by(model_outcome) %>%
    summarise(RSS = sum(resid, na.rm = TRUE),
              RMSE = sqrt(mean(resid^2, na.rm = TRUE))) %>% 
    rename(Outcome = model_outcome)
```

```{r fig-two, fig.asp = NULL, fig.width = 10, fig.height = 10, fig.out = "50%", fig.cap = "Observed values for each health outcome plotted against the posterior mean values (point) with their respective 95\\% posterior predictive interval (lines). A line of perfect prediction is shown. The points and lines are shown in red to indicate when the 95\\% posterior predictive interval lies outside the line of perfect prediction"}

fig_four_fev1 <- 
create_obs_pred_summary(jags_model_dt_summary_fev1,
                        jags_dat_model_fev1,
                        jags_dat_prep_fev1) %>%
    plot_fig_four() +
    xlim(20, 180) + 
    ylim(20, 180) +
    labs(title = "FEV1%",
         y = "Posterior Values") +
    theme(plot.title=element_text(size = 18))

fig_four_bmi <- 
create_obs_pred_summary(jags_model_dt_summary_bmi,
                        jags_dat_model_bmi,
                        jags_dat_prep_bmi) %>%
    plot_fig_four() +
    xlim(10, 110) + 
    ylim(10, 110) +
    labs(title = "BMI",
         y = "Posterior Values") +
    theme(plot.title=element_text(size = 18))

fig_four_sys <- 
create_obs_pred_summary(jags_model_dt_summary_sys,
                        jags_dat_model_sys,
                        jags_dat_prep_sys) %>%
    plot_fig_four() +
    xlim(60, 250) + 
    ylim(60, 250) +
    labs(title = "Systolic",
         y = "Posterior Values") +
    theme(plot.title=element_text(size = 18))

fig_four_dias <- 
create_obs_pred_summary(jags_model_dt_summary_dias,
                        jags_dat_model_dias,
                        jags_dat_prep_dias) %>%
    plot_fig_four() +
    xlim(20, 140) + 
    ylim(20, 140) +
    labs(title = "Diastolic",
         y = "Posterior Values") +
    theme(plot.title=element_text(size = 18))


gridExtra::grid.arrange(fig_four_fev1,
                        fig_four_bmi,
                        fig_four_sys,
                        fig_four_dias)

# figure 2 here used to be figure 5. Just for reference

```


**Model Fit**

Figure 2 shows the posterior predictions for each outcome plotted on the y axis
against the observed values on the x axis. The points represent the observed
values and the corresponding posterior means with vertical lines representing
the respective 95\% posterior predictive intervals. A line of perfect prediction
runs from the bottom left to the top right corner. The points and lines are
shown in red to indicate when the observed value lies outside of the 95\%
posterior predictive interval.

Model fit was assessed by visual inspection of Figure 2, and by assessing the percentage of observed values that lie within nominated posterior predictive intervals (Table 2). The models for BMI and FEV1\% had very high proportions of observed values in the 95\% intervals and 80\% intervals, indicating reasonable model fit.

```{r table-1-part-1}
 # if(interval == "95%"){

table_1_part_1 <- obs_pred_summary_all %>%
    group_by(model_outcome) %>%
    # obs_pred_summary %>%
        filter(forecasted == FALSE) %>%
        count(in_range_95) %>%
        na.omit() %>%
        mutate(pct = round((n / sum(n)), 4)) %>%
        mutate(pct = pct * 100) %>%
        mutate(n.paste = sprintf("%g (%0.2f%%)",n,  pct)) %>%
        select(model_outcome,
               in_range_95,
               pct) %>%
        spread(in_range_95, pct) %>%
        rename(`Inside 95% PI` = `inside`,
               `Outside PI` = `outside`) %>%
        select(-`Outside PI`) %>%
        ungroup() %>%
        rename(Outcome = model_outcome) %>%
        mutate(Outcome = plyr::revalue(Outcome,
                                       c(bmi="BMI",
                                         fev1="FEV1 (%)",
                                         dias="Diastolic BP",
                                         sys="Systolic BP")))

table_1_part_2 <-
    obs_pred_summary_all %>%
    group_by(model_outcome) %>%
            group_by(model_outcome) %>%
            filter(forecasted == FALSE) %>%
            count(in_range_80) %>%
            na.omit() %>%
            mutate(pct = round((n / sum(n)), 4)) %>%
            mutate(pct = pct * 100) %>%
            # mutate_each(funs(parse_number), n, pct) %>%
            mutate(n_paste = sprintf("%g (%0.2f%%)",n,  pct)) %>%
            select(model_outcome,
                   in_range_80,
                   pct) %>%
            spread(in_range_80, pct) %>%
            rename(`Inside 80% PI` = `inside`,
                   `Outside PI` = `outside`) %>%
            select(-`Outside PI`) %>%
            ungroup() %>%
            rename(Outcome = model_outcome) %>%
            mutate(Outcome = plyr::revalue(Outcome,
                                           c(bmi="BMI",
                                             fev1="FEV1 (%)",
                                             dias="Diastolic BP",
                                             sys="Systolic BP")))
    
```

```{r actually-print-table-1}
left_join(table_1_part_1,
          table_1_part_2,
          by = "Outcome") %>%
    knitr::kable(caption = "Percent of observed values inside the 95% and 80% posterior prediction intervals and RSS for each model outcome.")

```

```{r combine-summary-results}

jags_model_dt_summary_all <- 
    bind_rows(jags_model_dt_summary_fev1,
              jags_model_dt_summary_bmi,
              jags_model_dt_summary_sys,
              jags_model_dt_summary_dias)

```

```{r prepare-fig-four}

### beta_d ---------------------------------------------------------------------
jags_model_tidy_beta_d <- jags_model_dt_summary_all %>%
    filter(grepl("^beta_d\\[", parameter))

# there are 1804 unique individuals
# n_distinct(dat_isba_15$uin)
# so now we can look at those individuals who have an overall trend less than 0.

# but first we are going to do this for each of the datasets

uin_3_visits_bmi <- count(dat_isba_bmi,uin) %>%
    filter(n >= 3) %>%
    .$uin

uin_3_visits_fev1 <- count(dat_isba_fev1,uin) %>%
    filter(n >= 3) %>%
    .$uin

uin_3_visits_sys <- count(dat_isba_sys,uin) %>%
    filter(n >= 3) %>%
    .$uin

uin_3_visits_dias <- count(dat_isba_dias,uin) %>%
    filter(n >= 3) %>%
    .$uin

# then repeat this for bmi...sys.

id_more_than_3_bmi <- dat_isba_bmi %>% 
    mutate(which_uin = (uin %in% uin_3_visits_bmi)) %>%
    filter(which_uin == TRUE) %>%
    .$ID %>%
    unique()

id_more_than_3_fev1 <- dat_isba_fev1 %>% 
    mutate(which_uin = (uin %in% uin_3_visits_fev1)) %>%
    filter(which_uin == TRUE) %>%
    .$ID %>%
    unique()

id_more_than_3_sys <- dat_isba_sys %>% 
    mutate(which_uin = (uin %in% uin_3_visits_sys)) %>%
    filter(which_uin == TRUE) %>%
    .$ID %>%
    unique()

id_more_than_3_dias <- dat_isba_dias %>% 
    mutate(which_uin = (uin %in% uin_3_visits_dias)) %>%
    filter(which_uin == TRUE) %>%
    .$ID %>%
    unique()

beta_d_3_vec_bmi <- sprintf("beta_d[%s]",id_more_than_3_bmi)
beta_d_3_vec_fev1 <- sprintf("beta_d[%s]",id_more_than_3_fev1)
beta_d_3_vec_sys <- sprintf("beta_d[%s]",id_more_than_3_sys)
beta_d_3_vec_dias <- sprintf("beta_d[%s]",id_more_than_3_dias)


model_beta_d_3_vis_bmi <- jags_model_tidy_beta_d %>%
    filter(model_outcome == "bmi",
           parameter %in% beta_d_3_vec_bmi)

model_beta_d_3_vis_fev1 <- jags_model_tidy_beta_d %>%
    filter(model_outcome == "fev1",
           parameter %in% beta_d_3_vec_fev1)

model_beta_d_3_vis_sys <- jags_model_tidy_beta_d %>%
    filter(model_outcome == "sys",
           parameter %in% beta_d_3_vec_sys)

model_beta_d_3_vis_dias <- jags_model_tidy_beta_d %>%
    filter(model_outcome == "dias",
           parameter %in% beta_d_3_vec_dias)

bind_3_vis <- bind_rows(
    model_beta_d_3_vis_bmi,
    model_beta_d_3_vis_fev1,
    model_beta_d_3_vis_sys,
    model_beta_d_3_vis_dias
) %>%
        arrange(model_outcome) %>%
        group_by(model_outcome) %>%
        mutate(highlight = if_else(q_02.5 < 0 & q_975 < 0 |
                                       q_02.5 > 0 & q_975 > 0,
                                   true = "highlight",
                                   false = "lowlight")) %>%
               # Terms = param_name) %>%
        select(parameter,
            # Terms,
               q_02.5,
               mean,
               median,
               q_975,
               highlight,
               pr_less_zero, 
               model_outcome)

```

```{r fig-three, fig.height = 3, fig.width = 10, fig.cap = "Individuals whose parameter estimates for change over time, $\\beta_d$, were the furthest away from zero (and did not contain zero in the credible interval) for the health characteristics systolic blood pressure, FEV1%, diastolic blood pressure, and BMI."}


# alternative method, which looks at just taking the lowest value
library(magrittr) # get the %$% operator

worst_id_num <-
bind_3_vis %>%
    # get the id number out
    mutate(id_num = readr::parse_number(parameter)) %>%
    # get the mean absolute difference from zero
    mutate(abs_zero = abs(mean - 0)) %>%
    # then group by model outcome
    group_by(model_outcome) %>%
    # identify which id numbers are 
    filter(abs_zero == max(abs_zero)) %$%
    # use the special exposition operator from magrittr
    id_num

# I visually identified these individuals: 1512, 341, 324, 94

bind_3_vis %>%
    # get the id number out
    mutate(id_num = readr::parse_number(parameter)) %>%
    # only select those with these four numbers, which I 
    # have identified visually
    filter(id_num %in% worst_id_num) %>%
    mutate(scale_id_num = 
               if_else(id_num == 94,
                       1,
                       if_else(id_num == 324,
                               true = 2,
                               if_else(id_num == 341,
                                       3,
                                       if_else(id_num == 1512,
                                               4,
                                               0))))) %>%
    ggplot(aes(x = mean,
               y = model_outcome,
               colour = highlight)) +
        geom_point(size = 1.5) +
        geom_errorbarh(aes(xmin = q_02.5,
                           xmax = q_975),
                       height = 0.1,
                       size = 1) +
        labs(x = "",
             y = "") +
        geom_vline(xintercept = 0,
                   colour = "coral",
                   linetype = 2) +
        scale_colour_manual(values = c("darkblue", "gray")) +
        theme_minimal() +
        theme(legend.position = "none",
              axis.text.x= element_text(colour = "#003366",
                                        size = 30),
              axis.text.y= element_text(colour = "#003366",
                                        size = 24)) %>%
    scale_fill_gradient(low = "grey50", high = "darkblue") +
    facet_grid(~scale_id_num, 
               labeller = label_bquote(cols = paste(beta[d]))) +
    theme_bw() +
    theme(legend.position = "none", 
          axis.text.x= element_text(angle = 45,
                                    hjust = 1))

```

Figure 3 shows four selected "at risk" individuals and their posterior mean and
credible intervals for the health characteristics systolic blood pressure,
FEV1\%, diastolic blood pressure, and BMI. The proportion of individuals "at
risk" for each health outcome, and the mean and standard deviation for each
health outcome for those at risk and not at risk, are shown in Table 3.
Individuals were identified as "at risk" in this case according to whether their
parameter estimates for change over time $\beta_d$ were the furthermost away
from zero (and did not contain zero in the credible interval) for the health
characteristics systolic blood pressure, FEV1\%, diastolic blood pressure, and
BMI. As described in the method section, these individuals had 3 or more visits.

\newpage 

```{r print-table-2}
prop_in_out_data <- bind_3_vis %>%
    mutate(ID = readr::parse_number(parameter)) %>%
    select(ID, highlight) %>%
    spread(key = model_outcome,
           value = highlight,
           sep = "_")  %>%
    left_join(dat_isba_bmi,
              key = "ID") %>%
    select(ID,
           model_outcome_bmi,
           model_outcome_fev1,
           model_outcome_sys,
           model_outcome_dias,
           bmi,
           fev1_perc,
           sys,
           dias)

prop_in_risk_sum <- function(data,
                             ID = "ID",
                             outcome_col,
                             outcome){
    
    dat_sum <- data %>%
    select_(outcome_col,
            outcome) %>%
    group_by_(outcome_col) %>%
    summarise_if(is.numeric, 
                 funs(mean, sd), 
                 na.rm = TRUE) %>%
        mutate(summary = sprintf("%0.2f (%0.2f)",mean,  sd),
               # drop mean and sd
               mean = NULL, 
               sd = NULL) %>%
        # hard coded, because plyr::revalue is hard
        mutate(new_info = c("Mean (SD) at risk",
                            "Mean (SD) not at risk")) %>%
        select(new_info,
               summary)
    
    # NSE is hard
    names(dat_sum) <- c("summary",outcome)
    
    return(dat_sum)
}
    
table_2 <- 
prop_in_risk_sum(prop_in_out_data,
                 outcome_col = "model_outcome_bmi",
                 outcome = "bmi") %>%
    left_join({prop_in_risk_sum(prop_in_out_data,
                                outcome_col = "model_outcome_fev1",
                                outcome = "fev1_perc")},
              by = "summary") %>%
    left_join({prop_in_risk_sum(prop_in_out_data,
                 outcome_col = "model_outcome_sys",
                 outcome = "sys")},
              by = "summary") %>% 
    left_join({prop_in_risk_sum(prop_in_out_data,
                 outcome_col = "model_outcome_dias",
                 outcome = "dias")}) %>%
# also, this is the proportion of individuals who have three visits.
    bind_rows({bind_3_vis %>%
    group_by(model_outcome) %>%
            summarise(prop_inside = round(mean(highlight == "highlight"),2)) %>%
            mutate(prop_inside = as.character(prop_inside)) %>%
            spread(key = model_outcome,
                   value = prop_inside) %>%
            rename(fev1_perc = fev1) %>%
            mutate(summary = "Proportion at risk")}) %>%
    mutate(row_order = c(3,2,1)) %>%
    arrange(row_order) %>%
    select(-row_order)

names(table_2) <- c("Summary", 
                    "BMI",
                    "FEV1%",
                    "Systolic",
                    "Diastolic")

    knitr::kable(table_2,
                 caption = "The proportion of individuals with 3 or more visits who were 'at risk' (95% credible interval for change over time did not include zero) and the health outcomes for the 'at risk' and 'not at risk' groups")


```

Table 4 shows the posterior mean, 95% credible interval, and probability of being negative for each of the risk factors considered, namely smoking, dust, and days, since commencement.


```{r fig-two-summary-all, fig.width = 6, fig.height = 6}

# get all the beta single terms
# beta_all[1] is the intercept, which we are not interested in.
# keep in mind, that this is based on the idea that I am using the same
# model.matrix as I was 4 months ago.
# now, let's instead do something a bit more cleve, and look at the jags model
# list used, and look at the head of the matrix, and the names.

x_names <- colnames(jags_dat_model_bmi$X)
which_col_seg <- grep("^SEG",x_names)
full_vec <- 1:length(x_names)
which_col_not_seg <- full_vec[grepl("^SEG",x_names) == FALSE]

# find the names of these characters    
# these are gender, smoked, dust
beta_singles <- sprintf("beta_all[%s]", which_col_not_seg)

# look at parameter beta_dc - the effect of DAY.

jags_model_beta_day <- jags_model_dt_summary_all %>%
    filter(grepl("^beta_dc", parameter))
    

# get the parameters from the data
fig_two_data <- jags_model_dt_summary_all[parameter %in% beta_singles] %>%
    bind_rows(jags_model_beta_day) %>%
    arrange(model_outcome,parameter) %>%
    group_by(model_outcome) %>%
    mutate(term_names = c("gender",
                          "smoking",
                          "dust",
                          "days"),
           highlight = if_else(q_02.5 < 0 & q_975 < 0 |
                                   q_02.5 > 0 & q_975 > 0,
                               true = "highlight",
                               false = "lowlight")) %>%
    ungroup() %>%
    select(term_names,
           q_02.5,
           mean,
           median,
           q_975,
           highlight,
           pr_less_zero,
           everything())


```

```{r}

table_1 <- fig_two_data %>%
    select(term_names,
           q_02.5,
           mean,
           q_975,
           pr_less_zero,
           model_outcome) %>% 
    mutate(q_02.5 = round(q_02.5,2),
           mean = round(mean,2),
           q_975 = round(q_975,2),
           pr_less_zero = round(pr_less_zero,2)) %>%
    rename(Outcome = model_outcome,
           Terms = term_names,
           `L 95% CI` = q_02.5,
           `U 95% CI` = q_975,
           `P(Term < 0)` = pr_less_zero)
    
    # probability < 0 . if 0, it is 0 to 7 places.

table_1_bmi <- table_1 %>%
    filter(Outcome == "bmi") %>% 
    select(-Outcome)

table_1_fev1 <- table_1 %>% 
    filter(Outcome == "fev1") %>%
    select(-Outcome)

table_1_sys <- table_1 %>% 
    filter(Outcome == "sys") %>%
    select(-Outcome)

table_1_dias <- table_1 %>% 
    filter(Outcome == "dias") %>%
    select(-Outcome)

```

```{r background-creating-table-two}

my_format <- function(x) format(x,scientific = TRUE, digits = 3)

my_table_two_prep <- function(data){
    
    data %>%
    mutate(est = my_format(mean),
           cr_int = paste0(
                           "(",
                            my_format(`L 95% CI`),
                            ", ",
                            my_format(`U 95% CI`),
                            ")"),
           est_pr = paste0(my_format(`P(Term < 0)`))) 
    
}

super_table_two <-
fig_two_data %>%
    select(term_names,
           q_02.5,
           mean,
           q_975,
           pr_less_zero,
           model_outcome) %>% 
    rename(Outcome = model_outcome,
           Terms = term_names,
           `L 95% CI` = q_02.5,
           `U 95% CI` = q_975,
           `P(Term < 0)` = pr_less_zero)

super_table_two_days <- super_table_two %>% filter(Terms == "days")

super_table_two_not_days <- super_table_two %>% filter(Terms != "days")

super_table_two_not_days_summary_pt1 <- super_table_two_not_days %>% 
    mutate(est = paste0(round(mean,2)),
           cr_int = paste0("(",
                         round(`L 95% CI`,2),
                         ", ",
                         round(`U 95% CI`,2),
                         ")"),
           est_pr = round(`P(Term < 0)`,2)) %>%
    select(Terms,
           Outcome,
           est) %>%
    filter(Terms != "gender") %>%
    spread(key = Outcome,
           value = est)

super_table_two_not_days_summary_pt2 <- super_table_two_not_days %>% 
    mutate(est = round(mean,2),
           cr_int = paste0("(",
                         round(`L 95% CI`,2),
                         ", ",
                         round(`U 95% CI`,2),
                         ")"),
           est_pr = round(`P(Term < 0)`,2)) %>%
    select(Terms,
           Outcome,
           cr_int) %>%
    filter(Terms != "gender") %>%
    spread(key = Outcome,
           value = cr_int)


super_table_two_not_days_summary_pt3 <- super_table_two_not_days %>% 
    mutate(est = round(mean,2),
           cr_int = paste0("(",
                         round(`L 95% CI`,2),
                         ", ",
                         round(`U 95% CI`,2),
                         ")"),
           est_pr = paste(round(`P(Term < 0)`,2))) %>%
    select(Terms,
           Outcome,
           est_pr) %>%
    filter(Terms != "gender") %>%
    spread(key = Outcome,
           value = est_pr)

table_two_day_master <- my_table_two_prep(super_table_two_days)
    

new_summary_table_two <- bind_rows({
    table_two_day_master %>%
        select(Terms,
               Outcome,
               est) %>%
        spread(key = Outcome,
               value = est)
},{
        table_two_day_master %>%
        select(Terms,
               Outcome,
               cr_int) %>%
        spread(key = Outcome,
               value = cr_int)
    },{
        table_two_day_master %>%
        select(Terms,
               Outcome,
               est_pr) %>%
        spread(key = Outcome,
               value = est_pr)
    }) %>%
    bind_rows(super_table_two_not_days_summary_pt1,
              super_table_two_not_days_summary_pt2,
              super_table_two_not_days_summary_pt3) %>%
    arrange(Terms) %>%
    mutate(Terms = c("days",
                     "",
                     "",
                     "dust",
                     "",
                     "",
                     "smoking",
                     "",
                     "")) %>%
    select(Terms, 
           bmi,
           fev1,
           sys,
           dias)
    
names(new_summary_table_two) <- c("Terms",
                                  "BMI",
                                  "FEV1%",
                                  "Systolic",
                                  "Diastolic"
                                             )

```

\newpage


-------------------------------------------------------------------------------------------------------------------------------------
  Terms             BMI                                FEV1%                     Systolic                        Diastolic       
--------- ------------------------------ ------------------------------ ------------------------------ ------------------------------
  days       6.09 x 10^-4^                -8.44 x 10^-4^                 1.35 x 10^-3^                  1.78 x 10^-3^        

          (4.72x10^-4^, 7.52x10^-4^)      (-1.41x10^-3^, -2.51x10^-4^)   (7.30x10^-4^, 1.97x10^-3^)       (1.34x10^-3^, 2.17x10^-3^) 

                 0                         9.98 x 10^-1^                  0                                0        

  dust             0.11                     0.33                          0.24                            -0.21         

               (-0.06, 0.3)               (-0.52, 1.15)                   (-0.72, 1.16)                   (-0.92, 0.51)     

                   0.11                     0.22                          0.31                            0.73          

 smoking           -0.33                   -1.87                          0.47                            0.52          

               (-0.72, 0.07)              (-3.24, -0.55)                  (-0.8, 1.79)                    (-0.38, 1.44)     

                   0.94                      1                            0.24                            0.13          
-------------------------------------------------------------------------------------------------------------------------------------

Table: Estimated posterior mean, 95% credible intervals and probability of the effect of Day, Dust, and Smoking parameters being less than zero.

The number of days since first visit had a substantive effect on all outcomes,
and was associated with a decrease in FEV~1~\%, and a decrease in BMI,
diastolic and systolic blood pressure. Dust did not have a substantive impact on
any outcomes, but was associated with an 11\% chance of decreased BMI, a 22\%
chance of decreased FEV~1~\%, a 31\% chance of decreased systolic blood
pressure, and a 73\% chance of decreased diastolic blood pressure. Being a
smoker was associated with substantively decreased FEV~1~\%, a 100\% chance
of decreased FEV~1~ \%, a 95\% chance of decreased BMI, a 13\% chance of
increased diastolic blood pressure, and a 24\% chance of increased systolic
blood pressure.

```{r fig-four, fig.width = 8, fig.out = "50%", fig.height = 5, fig.cap = "Posterior mean and 95% credible interval for exposure group parameters, for each model. The baseline exposure group is Administration."}
# jags_model_dt_summary_all

x_names <- colnames(jags_dat_model_bmi$X)

which_col_seg <- grep("^SEG",x_names)

# which columns are for SEG?
which_col_seg <- grep("^SEG",x_names)

    beta_segs <- sprintf("beta_all[%s]", which_col_seg)
    
    fig_three_data <-
        jags_model_dt_summary_all[parameter %in% beta_segs] %>%
        mutate(param_num = readr::parse_number(parameter),
               param_name = param_labeller_seg(param_num, 
                                               which_col_seg),
               param_name_2 = label_seg_simple(param_num)) %>%
        arrange(model_outcome, param_num) %>%
        group_by(model_outcome) %>%
        mutate(highlight = if_else(q_02.5 < 0 & q_975 < 0 |
                                       q_02.5 > 0 & q_975 > 0,
                                   true = "highlight",
                                   false = "lowlight"),
               Terms = param_name_2) %>%
        select(Terms,
               q_02.5,
               mean,
               median,
               q_975,
               highlight,
               pr_less_zero, 
               model_outcome)

    
new_exposure_group <- c("Administration",
                        "Emergency",
                        "Exterior Maintenance",
                        "Technicians",
                        "Technology",
                        "Interior Maintenance",
                        "Field Experts")
  
fig_three_data %>%
    mutate(pr_less_zero = round(pr_less_zero, 2)) %>%
    # filter down the appropriate groups
    filter(Terms %in% new_exposure_group) %>%
    mutate(facet_names = case_when(
        model_outcome == "bmi" ~ "BMI",
        model_outcome == "dias" ~ "Diastolic Blood Pressure",
        model_outcome == "sys" ~ "Systolic Blood Pressure",
        model_outcome == "fev1" ~ "FEV1%"
    )) %>%
    ggplot(aes(x = mean,
               y = reorder(Terms, mean),
               colour = highlight)) +
        geom_point(size = 2) +
        geom_errorbarh(aes(xmin = q_02.5,
                           xmax = q_975),
                       height = 0.2,
                       size = 0.5) +
        labs(x = "",
             y = "") +
        geom_vline(xintercept = 0,
                   colour = "coral",
                   linetype = 2) +
        scale_colour_manual(values = c("darkblue", "gray")) +
        theme_minimal() +
        theme(legend.position = "none",
              axis.text.x= element_text(colour = "#003366",
                                        size = 30),
              axis.text.y= element_text(colour = "#003366",
                                        size = 24)) %>%
    scale_fill_gradient(low = "grey50", high = "darkblue") +
    facet_wrap(~facet_names) +
    theme_bw() +
    theme(legend.position = "none")

# If I want to preserve the order of the mean in each facet, I'll have to do something like this:
# https://stackoverflow.com/questions/24255305/varying-factor-order-in-each-facet-of-ggplot2

```

All exposure group effects were compared to the baseline Administration, and the effects for each outcome over all exposure groups are shown in Figure 4. BMI
was substantively lower in Maintenance, Emergency, and Field Experts. Diastolic
blood pressure was substantively lower in Technology and Field Expert exposure
groups. Technologists had substantively lower systolic blood pressure, and
Technicians had substantively higher FEV~1~\%.

```{r plot-figure-five,fig.height = 8, fig.width = 8}

# add a separate hline for bmi, dias, fev1, sys.

dat_hline <- tibble::tribble(
    ~model_outcome, ~value,
          "bmi",      30,
          "dias",     90,
          "fev1",     70,
          "sys",      140
)

# jags_model_dt_summary_y
fig_five_data_forecasted <-
    obs_pred_summary_all %>%
    mutate(last_observed = lead(forecasted),
           last_two = last_observed + forecasted) %>%
    filter(last_two == 1) %>%
    mutate(pr_chronic_disease = round(pr_chronic_disease,4))

# this is the complement to the previous dataframe
fig_five_data_observed <- 
    obs_pred_summary_all %>%
    filter(forecasted == FALSE)

```

Figure 5 shows the same four selected individuals previously identified as being
"at risk", from Figure 3, and the observed and predicted values for health
outcomes. Observed outcomes are shown as blue points and model mean posterior
values are shown as a blue line. The dark blue ribbon around the blue line
represents an 80\% credible interval, and the light blue ribbon the 95\% credible
interval. A one year forecast is shown as a red line extending from the blue
line, and similarly the 80\% and 95\% credible intervals are displayed. A dotted
line is shown for each outcome, which represents a clinically relevant threshold
of chronic disease for each outcome. Individual probabilities of chronic condition
in the one year forecast are labelled directly on Figure 5.


```{r fig-five, fig.height = 8, fig.width = 10, fig.out = "80%", fig.cap = "Individuals from Figure 3 and their observed (points), predicted (blue region and line), and forecasted (red region and lines) values for the health outcomes shown with 80% (darker region) and 95% credible intervals (lighter region). Labels show the probability of the individual having a clinically defined chronic disease at the forecasted timepoint."}

paste_3_letters <- function() paste0(sample(LETTERS,3), collapse = "")

id_94 <- sprintf("194-%s", paste_3_letters())
id_324 <- sprintf("1324-%s", paste_3_letters())
id_341 <- sprintf("1341-%s", paste_3_letters())
id_1512 <- sprintf("11512-%s", paste_3_letters())

fig_five_plot_data <- fig_five_data_observed %>%
    filter(ID %in% worst_id_num) %>%
    mutate(ID = case_when(
        ID == 94 ~ id_94,
        ID == 324 ~ id_324,
        ID == 341 ~ id_341,
        ID == 1512 ~ id_1512
    ))


fig_five_plot_data_forecast <- fig_five_data_forecasted %>%
    filter(ID %in% worst_id_num) %>%
    mutate(ID = case_when(
        ID == 94 ~ id_94,
        ID == 324 ~ id_324,
        ID == 341 ~ id_341,
        ID == 1512 ~ id_1512
    ))

ggplot(fig_five_plot_data,
           aes(x = days_since_arrival,
               y = pred_mean,
               group = ID)) +
    geom_line(colour = "steelblue") +
    geom_ribbon(aes(ymin = pred_025,
                    ymax = pred_975),
                alpha = 0.25,
                fill = "steelblue") +
    geom_ribbon(aes(ymin = pred_010,
                    ymax = pred_90),
                alpha = 0.25,
                fill = "navyblue") +
    geom_point(aes(y = outcome),
               colour = "steelblue") +
    geom_line(data = fig_five_plot_data_forecast,
              aes(x = days_since_arrival,
                  y = pred_mean),
              colour = "coral",
              linetype = 1) +
    geom_ribbon(data = fig_five_plot_data_forecast,
                aes(ymin = pred_025,
                    ymax = pred_975),
                alpha = 0.25,
                fill = "coral") +
    geom_ribbon(data = fig_five_plot_data_forecast,
                aes(ymin = pred_010,
                    ymax = pred_90),
                alpha = 0.25,
                fill = "red") +
    theme_bw() +
    # facet_wrap(~ID) + 
    labs(x = "Days Since Arrival",
         y = "Outcome") +
    facet_grid(model_outcome~ID,
               scales = "free_y",
               switch = "y") +
    geom_hline(data = dat_hline,
               aes(yintercept = value),
               colour = "red",
               linetype = 3) + 
    geom_label(data = filter(fig_five_plot_data_forecast, forecasted == TRUE),       
               aes(label = round(pr_chronic_disease,2),
                   vjust = 2),
               colour = "white",
               fill = "salmon")
    
```



```{r table-of-pr-chronic-disease}

forecasted_chronic_disease <- fig_five_data_forecasted %>% 
    filter(ID %in% worst_id_num,
           forecasted == TRUE) %>%
    select(model_outcome,
           ID,
           pr_chronic_disease) %>%
    mutate(pr_chronic_disease = round(pr_chronic_disease,4)) %>%
    spread(key = model_outcome,
           value = pr_chronic_disease) 

```

# Discussion

This paper set out to develop a Bayesian approach to analysing OHS data and to
illustrate three analytic aims focussed on the health outcomes lung function
(FEV~1~\%), Body Mass Index (BMI), and systolic and diastolic blood pressure.
Aim 1 investigated patterns and trends in the health outcomes. Aim 2 explored
the effects of smoking and industrial exposures over time for individuals and
worker groups. Aim 3 identified future individual risk of chronic conditions.

Aim 1 was addressed by examining individual change over time and identifying
individuals with the greatest degree of change over time for each outcome. We
demonstrated how one could then assess the overall change over time for these
individuals, which could be used to identify trends in other health outcomes. We
also identified the proportion of individuals who fell into an "at risk"
category. Identifying those individuals with substantive negative change in BMI,
lung function, systolic, and diastolic blood pressure means that medical
professionals could flag these individuals as at risk (compared to the overall
worker population), and provide more frequent medical attention to better
monitor their health.

Aim 2 was addressed by examining the probability that the effects of smoking and
dust were different from zero for each outcome, finding those exposure groups
that were substantively different from the reference group (Administration), and
identifying individuals at risk based on substantive change over time in health
outcomes. Smoking was associated with negative health outcomes for lung function
and systolic and diastolic blood pressure. This information can be used to
further support health policies, such as implementation of tobacco bans in the
workplace.

The nominated industrial exposure, dust was not substantively associated with
health outcomes in the workplaces in this case study. The relevant parameter
estimates had quite wide credible intervals, possibly due to interpolation of
the data, which was used to align dust measurement points with health
measurements. This demonstrates that frequent measurements of industrial
exposures of concern can provide more certainty in the measurement of effects.
Interestingly, results identified that employees in Administration should be
more closely monitored and perhaps should be the focus of health interventions
and healthy worker programs in workplaces. Providing descriptive statistics
of the outcomes for at risk and not at risk populations (Table 3) provides
medical professionals with a measure of how meaningfully different these
populations are, and facilitates more targeted health and wellness programs.

Aim 3 was addressed by calculating posterior predictions and corresponding
intervals and one year forecasts for all individuals. This allows medical
professionals to assign a probability that an individual might move into an "at
risk" category one year from their last visit. This means that individuals may
be flagged as "at risk" and further action can be taken, perhaps in the form of
more frequent medical visits to more closely monitor their health measures. his
demonstrates how forecasting could identify "at risk" individuals, by placing a
threshold on the probability of health outcomes being medically classified as
chronic or acute conditions. Some individuals might cross the threshold over
the observed times, whilst others might be predicted to enter the threshold with
a given probability over the next year.

The Bayesian hierarchical model accounts for important features of data such as
multiple measurements for individuals, and the exposure group structure in the 
workplace. The definition of a Bayesian credible interval as a
range of probable values for a parameter makes it easier to communicate model
inferences. The model also provides probabilities of interest directly,
conditional on the data. This is a useful complement to credible intervals.
Forecasting of future observations in a Bayesian framework also allows for
probabilistic statements based directly on the posterior predictive
distribution. The Bayesian framework naturally includes additional
uncertainty due to imputation of missing values. These features compare
favourably to their frequentist modelling counterparts.

Extensions to the Bayesian models developed here are also straightforward. For
example, an obvious next step in analysis might be to add interactions into the model, such as smoking and dust, or BMI and blood pressure. One relatively straightforward way to explore the impact of interactions is to evaluate the Bayes factors for each variable, approximated using the Savage-Dickey density ratio, which only requires samples from the posterior [@Wagenmakers2010 ; Dickey1970]. This can add time to the model building process, in terms of deciding upon the most useful model, but is worth the effort if the practitioner is genuinely interested in one or two interaction terms. As with any working population, there may be some healthy cohort effect
[@Li1999; @McMichael1974], where, being employable, employees are healthier than
the general population.  The methods provided in this study identify employees
and groups that are different from the population. Combining this information
with reference chronic conditions provides a more comprehensive approach which
might otherwise have missed healthy employees.

It is also possible that there may perhaps be
less measurement error for long term employees; here a model that predicts the
number of visits for each individual may be useful, where the number of visits
$n_{0i}$ for each individual is the outcome. Additionally, there may be some
correlation between the slope and the intercept, which could be accounted for by
modelling them as coming from some bivariate normal distribution [@BDA3]. It is acknowledged that the model fit is not ideal, particularly with respect to underestimation and overestimation of very high and low values, respectively. While this regression to the mean is to be expected given the random effects terms, the fit could be improved for the other outcomes, perhaps by including interactions as previously discussed.

As far as we are aware, this is the first time Bayesian methods have been
applied to this kind of OHS data. It is our hope that this paper can serve as
one way to fit and interpret these data, and serve as encouragement for researchers in
the field of OHS, to include Bayesian approaches in
their analytic toolkit. The ultimate ambition is to provide more informative
evidence-based OHS assessments for a healthier workforce and more profitable
workplaces.

## Acknowledgements

The authors would like to thank Dr. Xing Lee and Dr. Nicole White for their advice
and collaboration with this project.


# References
